---
title: "Assignment_8"
author: "Weijie Gao"
date: "11/28/2017"
output:
  word_document: default
  html_document: default
---

```{r}
suppressWarnings(library(forecast))
suppressWarnings(library(xts))
suppressWarnings(library(fGarch))
```

```{r}
dataPath <- "~/Google Drive/2017 Fall/Time Series/Week8"
seat_price <- read.csv(paste(dataPath,"seat_price.csv",sep='/'), header=TRUE)
classification_data <- read.csv(paste(dataPath,"Contracts_Classification.csv",sep='/'), header=TRUE)
volume_data <- read.csv(paste(dataPath,"Contracts_Volume.csv",sep='/'), header=TRUE)
```


```{r}
head(seat_price)
```

```{r}
head(classification_data)
```

```{r}
head(volume_data)
```

#### 1. CME
#### Task A
```{r}
seat <- "CME"
commodity <- c(as.character(classification_data$Commodity.Code[classification_data$Division=="CME"]))
commodity <- unique(commodity)
commodity
```

```{r}
tradable_commodity <- is.element(volume_data$Commodity.Indicator,commodity)
head(tradable_commodity,3)
```

```{r}
volume <- volume_data[tradable_commodity,]
head(volume,3)
```


```{r}
volume$Electronic.Volume <- as.numeric(gsub(",","",volume$Electronic.Volume,fixed=TRUE))
volume$Floor.Volume <- volume$Total.Volume-volume$Electronic.Volume
head(volume,3)
```

```{r}
# aggregate upon derivative types.
cme_volume_0 <- aggregate(cbind(Electronic.Volume, Floor.Volume,
Total.Volume)~Date+Commodity.Indicator, data=volume, sum)
head(cme_volume_0,3)
```

```{r}
dim(cme_volume_0)
```

```{r}
# drop any data before date 1/1/2001
cme_volume <- cme_volume_0[!as.Date(cme_volume_0$Date, "%m/%d/%Y") < as.Date("2001-01-01"),]
head(cme_volume,3)
```

```{r}
dim(cme_volume)
```

```{r}
names(cme_volume)
```

```{r}
# aggregate CME based on Date
cme_volume<-aggregate(cbind(Electronic.Volume, Floor.Volume, Total.Volume)~Date, data=cme_volume,sum)
head(cme_volume,3)
```

```{r}
# order by date
cme_volume <- cme_volume[order(as.Date(cme_volume$Date, "%m/%d/%Y")),]
head(cme_volume,5)
```

```{r}
tail(cme_volume,5)
```

```{r}
# combine trading volume with seat price.
cme_seat_price <- seat_price[[toupper(seat)]]
head(cme_seat_price)
```

```{r}
cme_volume$Seat.Price <- cme_seat_price
head(cme_volume,3)
```


#### Split data into train and test
```{r}
# Partion the training and testing data based on the date of trading.
# training data: data with date before year 2013
# testing data: data with date in year 2013
cme_train <- cme_volume[as.Date(cme_volume$Date,"%m/%d/%Y") < as.Date("2013-01-01") ,]
cme_test <- cme_volume[!as.Date(cme_volume$Date,"%m/%d/%Y") < as.Date("2013-01-01") ,]
# create an empty list to store forecast values
seat_price_forecast <- list()
```

#### 1.1 Linear regression
```{r}
cme_lm <- lm(Seat.Price~Electronic.Volume+Floor.Volume, data=cme_train)
summary(cme_lm)
```

```{r}
seat_price_forecast$lm <- predict(cme_lm, cme_test)
```

#### 1.2 Linear regression with ARMA errors (use arima with xreg)
```{r}
cme_lm_arma_errors <- auto.arima(cme_train$Seat.Price, xreg=cme_train[,c(2,3)], allowdrift = FALSE)
summary(cme_lm_arma_errors)

seat_price_forecast$lm_arma_errors <- predict(cme_lm_arma_errors, n.ahead=12, newxreg = cme_test[,c(2,3)])$mean
```

#### 1.3 ARIMA
```{r}
cme_arima <- auto.arima(cme_train$Seat.Price, allowdrift = FALSE)
summary(cme_arima)
```

```{r}
seat_price_forecast$arima <- forecast(cme_arima,h=12)$mean
```

#### 1.4 Seasonal ARIMA (SARIMA)
```{r}
cme_sarima <- auto.arima(ts(cme_train$Seat.Price,frequency = 12), allowdrift = FALSE)
summary(cme_sarima)
```

#### 1.5 Fractional ARIMA (ARFIMA)
```{r}
acf(cme_train$Seat.Price, main = "CME Seat Price Before 2013")
cme_arfima <- arfima(cme_train$Seat.Price)
summary(cme_arfima)
seat_price_forecast$arfima <- forecast(cme_arfima , h=12)$mean
```

#### 1.6 ARMA and GARCH combination
```{r}
cme_seat_price_xts <- xts(cme_train$Seat.Price, order.by = as.Date(cme_train$Date, "%m/%d/%Y"))
cme_log <- log(cme_seat_price_xts)
# in order to make data stationary, need to take first difference of cme_log
cme_log <- diff(cme_log)[-1]
plot(cme_log)
```

```{r}
# find p,q
cme_arma <- auto.arima(cme_log)
summary(cme_arma)
```

```{r}
# from summary results, we will choose p=2, q=2
cme_garch <- garchFit(~arma(2,2) + garch(1,1), data=cme_log, cond.dist = "std", trace=F)
summary(cme_garch)
```

```{r}
# forecast value
forecast_cme_log_garch <- predict(cme_garch, n.ahead=12)$meanForecast
# compute the price from forecast_cme_log_garch
seat_price_forecast$garch <- as.numeric(tail(cme_seat_price_xts,1)*exp(cumsum(forecast_cme_log_garch)))
```

#### Task B (sMAPE)
```{r}
smape <- function(fitted, actual) {
          return(2*mean(abs(fitted - actual) / (abs(fitted) + abs(actual))))
          }
cme_smape <- mapply(smape, seat_price_forecast, list(cme_test$Seat.Price))
cme_smape
```

```{r}
cme_smape[which.min(cme_smape)]
```
From sMAPE results, the Seasonal ARIMA model is the best one to forecast monthly prices for CME seat
classes since SARIMA model has the smallest sMAPE thus the best.

#### 2. IMM

#### Task A
```{r}
# IMM commodity subset
seat <- "IMM"
commodity_imm <- c(as.character(classification_data$Commodity.Code[classification_data$Division=="IMM"]))
commodity_imm <- unique(commodity_imm)
commodity_imm
```

```{r}
# filter out tradable commodities
tradable_commodity_imm <- is.element(volume_data$Commodity.Indicator,commodity_imm)
head(tradable_commodity_imm,3)
```

```{r}
volume_imm <- volume_data[tradable_commodity_imm,]
head(volume_imm,3)
```

```{r}
# floor volume
volume_imm$Electronic.Volume <- as.numeric(gsub(",","",volume_imm$Electronic.Volume,fixed=TRUE))
volume_imm$Floor.Volume <- volume_imm$Total.Volume-volume_imm$Electronic.Volume
head(volume_imm,3)
```

```{r}
# aggregate upon derivative types.
imm_volume_0 <- aggregate(cbind(Electronic.Volume, Floor.Volume,
Total.Volume)~Date+Commodity.Indicator, data=volume_imm, sum)
head(imm_volume_0,3)
```

```{r}
dim(imm_volume_0)
```

```{r}
# drop any data before date 1/1/2001
imm_volume <- imm_volume_0[!as.Date(imm_volume_0$Date, "%m/%d/%Y") < as.Date("2001-01-01"),]
head(imm_volume,3)
```

```{r}
dim(imm_volume)
```

```{r}
# aggregate IMM based on Date
imm_volume<-aggregate(cbind(Electronic.Volume, Floor.Volume, Total.Volume)~Date, data=imm_volume,sum)
head(imm_volume,3)
```

```{r}
# order by date
imm_volume <- imm_volume[order(as.Date(imm_volume$Date, "%m/%d/%Y")),]
head(imm_volume,3)
```

```{r}
# combine trading volume with seat price.
imm_seat_price <- seat_price[[toupper(seat)]]
head(imm_seat_price)
```

```{r}
imm_volume$Seat.Price <- imm_seat_price
head(imm_volume,3)
```

#### Split data into train and test
```{r}
imm_train <- imm_volume[as.Date(imm_volume$Date,"%m/%d/%Y") < as.Date("2013-01-01") ,]
imm_test <- imm_volume[!as.Date(imm_volume$Date,"%m/%d/%Y") < as.Date("2013-01-01") ,]
# create an empty list to store forecast values
seat_price_forecast_imm <- list()
```

#### 2.1 Linear regression
```{r}
imm_lm <- lm(Seat.Price~Electronic.Volume+Floor.Volume, data=imm_train)
summary(imm_lm)
```

#### 2.2 Linear regression with ARMA errors (use arima with xreg)
```{r}
imm_lm_arma_errors <- auto.arima(imm_train$Seat.Price, xreg=imm_train[,c(2,3)], allowdrift = FALSE)
summary(imm_lm_arma_errors)
```

```{r}
seat_price_forecast_imm$lm_arma_errors <- predict(imm_lm_arma_errors, n.ahead=12, newxreg = imm_test[,c(2,3)])$mean
```

#### 2.3 ARIMA
```{r}
imm_arima <- auto.arima(imm_train$Seat.Price, allowdrift = FALSE)
summary(imm_arima)
seat_price_forecast_imm$arima <- forecast(imm_arima,h=12)$mean
```

#### 2.4 Seasonal ARIMA (SARIMA)
```{r}
imm_sarima <- auto.arima(ts(imm_train$Seat.Price,frequency = 12), allowdrift = FALSE)
summary(imm_sarima)
seat_price_forecast_imm$sarima <- forecast(imm_sarima , h=12)$mean
```

#### 2.5 Fractional ARIMA (ARFIMA)
```{r}
acf(imm_train$Seat.Price, main = "IMM Seat Price Before 2013")
```

```{r}
imm_arfima <- arfima(imm_train$Seat.Price)
summary(imm_arfima)
seat_price_forecast_imm$arfima <- forecast(imm_arfima, h=12)$mean
```

#### 2.6 ARMA and GARCH combination
```{r}
imm_seat_price_xts <- xts(imm_train$Seat.Price, order.by = as.Date(imm_train$Date, "%m/%d/%Y"))
imm_log <- log(imm_seat_price_xts)
```

```{r}
# in order to make data stationary, need to take first difference of imm_log
imm_log <- diff(imm_log)[-1]
plot(imm_log)
```

```{r}
# find p,q
imm_arma <- auto.arima(imm_log)
summary(imm_arma)
```

```{r}
# from summary results, we will choose p=0, q=2
imm_garch <- garchFit(~arma(0,2) + garch(1,1), data=imm_log, cond.dist = "std", trace=F)
summary(imm_garch)
```

```{r}
# forecast value
forecast_imm_log_garch <- predict(imm_garch, n.ahead=12)$meanForecast
# compute the price from forecast_cme_log_garch

seat_price_forecast_imm$garch <- as.numeric(tail(imm_seat_price_xts,1)*exp(cumsum(forecast_imm_log_garch)))
```

#### Task B
```{r}
smape <- function(fitted, actual) {
          return(2*mean(abs(fitted - actual) / (abs(fitted) + abs(actual))))
          }
imm_smape <- mapply(smape, seat_price_forecast_imm, list(imm_test$Seat.Price))
imm_smape
```

```{r}
imm_smape[which.min(imm_smape)]
```

From sMAPE results, the Fractional ARIMA model is the best one to forecast monthly prices for IMM seat
classes since ARFIMA model has the smallest sMAPE thus the best. 

#### 3. IOM 
#### Task A 
```{r}
# IOM commodity subset
seat <- "IOM"
commodity_iom <- c(as.character(classification_data$Commodity.Code[classification_data$Division=="IOM"]))
commodity_iom <- unique(commodity_iom)
head(commodity_iom,30)
```

```{r}
# filter out tradable commodities
tradable_commodity_iom <- is.element(volume_data$Commodity.Indicator,commodity_iom)
head(tradable_commodity_iom,3)
```

```{r}
volume_iom <- volume_data[tradable_commodity_iom,]
head(volume_iom,3)
```

```{r}
# floor volume
volume_iom$Electronic.Volume <- as.numeric(gsub(",","",volume_iom$Electronic.Volume,fixed=TRUE))
volume_iom$Floor.Volume <- volume_iom$Total.Volume-volume_iom$Electronic.Volume
head(volume_iom,3)
```

```{r}
# aggregate upon derivative types.
iom_volume_0 <- aggregate(cbind(Electronic.Volume, Floor.Volume,
Total.Volume)~Date+Commodity.Indicator, data=volume_iom, sum)
head(iom_volume_0,3)
```

```{r}
# drop any data before date 1/1/2001
iom_volume <- iom_volume_0[!as.Date(iom_volume_0$Date, "%m/%d/%Y") < as.Date("2001-01-01"),]
head(iom_volume,3)
```

```{r}
# aggregate IOM based on Date
iom_volume<-aggregate(cbind(Electronic.Volume, Floor.Volume, Total.Volume)~Date, data=iom_volume,sum)
head(iom_volume,3)
```

```{r}
# order by date
iom_volume <- iom_volume[order(as.Date(iom_volume$Date, "%m/%d/%Y")),]
head(iom_volume,3)
```

```{r}
# combine trading volume with seat price.
iom_seat_price <- seat_price[[toupper(seat)]]
head(iom_seat_price)

iom_volume$Seat.Price <- iom_seat_price
head(iom_volume,3)
```

#### Split data into train and test
```{r}
# Partion the training and testing data based on the date of trading.
# training data: data with date before year 2013
# testing data: data with date in year 2013
iom_train <- iom_volume[as.Date(iom_volume$Date,"%m/%d/%Y") < as.Date("2013-01-01") ,]
iom_test <- iom_volume[!as.Date(iom_volume$Date,"%m/%d/%Y") < as.Date("2013-01-01") ,]
# create an empty list to store forecast values
seat_price_forecast_iom <- list()
```

#### 3.1 Linear regression
```{r}
iom_lm <- lm(Seat.Price~Electronic.Volume+Floor.Volume, data=iom_train)
summary(iom_lm)
seat_price_forecast_iom$lm <- predict(iom_lm, iom_test)
```

#### 3.2 Linear regression with ARMA errors (use arima with xreg)
```{r}
iom_lm_arma_errors <- auto.arima(iom_train$Seat.Price, xreg=iom_train[,c(2,3)], allowdrift = FALSE)
summary(iom_lm_arma_errors)
```

#### 3.3 ARIMA
```{r}
iom_arima <- auto.arima(iom_train$Seat.Price, allowdrift = FALSE)
summary(iom_arima)
```

```{r}
seat_price_forecast_iom$arima <- forecast(iom_arima,h=12)$mean
```

#### 3.4 Seasonal ARIMA (SARIMA)
```{r}
iom_sarima <- auto.arima(ts(iom_train$Seat.Price,frequency = 12), allowdrift = FALSE)
summary(iom_sarima)
```

```{r}
seat_price_forecast_iom$sarima <- forecast(iom_sarima , h=12)$mean
```

#### 3.5 Fractional ARIMA (ARFIMA)
```{r}
acf(iom_train$Seat.Price, main = "IOM Seat Price Before 2013")
iom_arfima <- arfima(iom_train$Seat.Price)
summary(iom_arfima)
seat_price_forecast_iom$arfima <- forecast(iom_arfima, h=12)$mean
```

#### 3.6 ARMA and GARCH combination
```{r}
iom_seat_price_xts <- xts(iom_train$Seat.Price, order.by = as.Date(iom_train$Date, "%m/%d/%Y"))
iom_log <- log(iom_seat_price_xts)
# in order to make data stationary, need to take first difference of iom_log
iom_log <- diff(iom_log)[-1]
plot(iom_log)
```

```{r}
# find p,q
iom_arma <- auto.arima(iom_log)
summary(iom_arma)
```

```{r}
# from summary results, we will choose p=2, q=0
iom_garch <- garchFit(~arma(2,0) + garch(1,1), data=iom_log, cond.dist = "std", trace=F)
summary(iom_garch)
```

#### Task B (sMAPE)
```{r}
smape <- function(fitted, actual) {
          return(2*mean(abs(fitted - actual) / (abs(fitted) + abs(actual))))
        }
iom_smape <- mapply(smape, seat_price_forecast_iom, list(iom_test$Seat.Price))
iom_smape
```

```{r}
iom_smape[which.min(iom_smape)]
```
From sMAPE results, the ARMA and GARCH combination model is the best one to forecast monthly prices
for IOM seat classes since the ARMA and GARCH combination model has the smallest sMAPE thus the
best.

