---
title: "Assignment_7"
author: "Weijie Gao"
date: "11/19/2017"
output: html_document
---

```{r}
suppressWarnings(library("TSA"))
suppressWarnings(library(xts))
suppressWarnings(library(reshape))
suppressWarnings(library(forecast))
suppressWarnings(library(tseries))
```

```{r}
dataPath <- "~/Google Drive/2017 Fall/Time Series/Week7"
cme <- read.csv(paste(dataPath,"cmeS.csv",sep='/'),header=TRUE)
imm <- read.csv(paste(dataPath,"immS.csv",sep='/'),header=TRUE)
iom <- read.csv(paste(dataPath,"iomS.csv",sep='/'),header=TRUE)
```

```{r}
head(cme)
```
```{r}
head(imm)
```

```{r}
head(iom)
```

```{r}
par(mfrow=c(3,1))
plot(cme[,4], type = "l", main = "CME", ylab = "Price")
plot(imm[,4], type = "l", main = "IMM", ylab = "Price")
plot(iom[,4], type = "l", main = "IOM", ylab = "Price")
```
```{r}
index <- 2001:2013
year <- c(rep(index,each=12))
month <- rep(1:12,13)
price <- rep(NA,length(month))
data <- as.data.frame(cbind(year,month,price))
head(data,5)
```

```{r}
dim(data)
data.cme <- data
data.imm <- data
data.iom <- data
```

```{r}
time <- melt(cme,id.vars = c("Year","Month"),
measure.vars = c("price"))
cme.new <- cast(time, Month+Year~variable, mean)
head(cme.new)
```

```{r}
dim(data.cme)
dim(cme.new)

for (i in 1:156){
  for (j in 1:134)
    if (data.cme[i,1] == cme.new[j,2] & data.cme[i,2] == cme.new[j,1])
    {
      data.cme[i,3] <- cme.new[j,3]
    }
}

head(data.cme)
```

For any months with missing seat price value, I will take average of the previous seat
prices and following seat prices as the seat price for that month.

```{r}
for (i in 2:155){
  if (is.na(data.cme[i,3]) == T){
    data.cme[i,3] <- (data.cme[i-1,3] + data.cme[i+1,3])/2
    }
}
head(data.cme)
```

```{r}
for (i in 7:10){
data.cme[i,3] <- (as.numeric(data.cme[6,3])+as.numeric(data.cme[11,3]))/2
}
head(data.cme)
```

```{r}
data.cme$year <- as.factor(data.cme$year)
data.cme$month <- as.factor(data.cme$month)

data.cme.ts <- ts(data.cme$price,start=2001,freq=12)
plot(data.cme.ts,main="time series plot of CME seat price")
```
```{r}
acf(data.cme.ts,main="ACF plot of CME seat price")
```

```{r}
pacf(data.cme.ts,main="PACF plot of CME seat price")
```

#### IMM
```{r}
time <- melt(imm,id.vars = c("Year","Month"),
measure.vars = c("price"))
imm.new <- cast(time, Month+Year~variable, mean)
head(imm.new)
```

```{r}
for (i in 1:156){
  for (j in 1:146)
    if (data.imm[i,1] == imm.new[j,2] & data.imm[i,2] == imm.new[j,1]){
    data.imm[i,3] <- imm.new[j,3]
    }
}
head(data.imm)
```

```{r}
for (i in 2:155){
  if (is.na(data.imm[i,3]) == T){
    data.imm[i,3] <- (data.imm[i-1,3] + data.imm[i+1,3])/2
  }
}

# check any rows = NA
which(is.na(data.imm[,3]) == T)
```

```{r}
for (i in 21:22){
  data.imm[i,3] <- (as.numeric(data.imm[20,3])+as.numeric(data.imm[23,3]))/2
}
```

```{r}
data.imm$year <- as.factor(data.imm$year)
data.imm$month <- as.factor(data.imm$month)


# plot dadta
data.imm.ts <- ts(data.imm$price,start=2001,freq=12)
plot(data.imm.ts,main="time series plot of IMM seat price")
```

```{r}
acf(data.imm.ts,main="ACF plot of IMM seat price")
```

```{r}
pacf(data.imm.ts,main="PACF plot of IMM seat price")
```

#### IOM
```{r}
time <- melt(iom,id.vars = c("Year","Month"),
measure.vars = c("price"))
iom.new <- cast(time, Month+Year~variable, mean)
iom.new <- iom.new[-1,]
head(iom.new)
```

```{r}
for (i in 1:156){
  for (j in 1:146)
    if (data.iom[i,1] == iom.new[j,2] & data.iom[i,2] == iom.new[j,1]){
    data.iom[i,3] <- iom.new[j,3]
  }
}
```

```{r}
for (i in 2:155){
  if (is.na(data.iom[i,3]) == T){
    data.iom[i,3] <- (data.iom[i-1,3] + data.iom[i+1,3])/2
  }
}
```

```{r}
# check any rows = NA
which(is.na(data.iom[,3]) == T)
```

```{r}
data.iom$year <- as.factor(data.iom$year)
data.iom$month <- as.factor(data.iom$month)

# plot data
data.iom.ts <- ts(data.iom$price,start=2001,freq=12)
plot(data.iom.ts,main="time series plot of IOM seat price")
```

```{r}
acf(data.iom.ts,main="ACF plot of IOM seat price")
```

```{r}
pacf(data.iom.ts,main="PACF plot of IOM seat price")
```

```{r}
write.csv(data.frame(Date=paste(data.cme$year,data.cme$month,sep="/"), CME=data.cme$price, IMM=data.imm$price,IOM=data.iom$price), file="seat_price.csv", row.names=FALSE)
```

This project is about missing value and to smoothen time series data. I used mean for interpolation because for months with multiple values, I think mean of seat prices is a good measure to reflect the average behavior of seat price in that month. Alternative methods such as median or modes is also considered, but given sample size is small and there is no obvious outlier, I suppose mean is acceptable. And in order to with missing values, I took averages of seat prices before and next to that month as this could be a good measure in reflecting seat behavior in this case.
